Vibrations and Heat Diffusion on the Unit Disc

### 2020-04-18 writeup/draft2.tex
Update draft2.tex
added conclusion

\documentclass{article}
\usepackage{changepage}
\usepackage{noah}

@@

@@

% \begin{center}
%   \begin{figure}
%     \includegraphics[width = 4.5 cm]{images/initialvalue.png}
%     \caption{Initial distribution}
%   \end{figure}
% \end{center}

@@

% \begin{figure}[!htb]
%   \minipage{0.32\textwidth}
%     \includegraphics[width=\linewidth]{images/heat1.png}
%   \endminipage\hfill
%   \minipage{0.32\textwidth}
%     \includegraphics[width=\linewidth]{images/heat2.png}
%   \endminipage\hfill
%   \minipage{0.32\textwidth}%
%     \includegraphics[width=\linewidth]{images/heat3.png}
%   \endminipage
%   \caption{Solution to the heat equation for various points in time.}
% \end{figure}

@@

@@

% \begin{figure}[!htb]
%   \minipage{0.32\textwidth}
%     \includegraphics[width=\linewidth]{images/wave1.png}
%   \endminipage\hfill
%   \minipage{0.32\textwidth}
%     \includegraphics[width=\linewidth]{images/wave2.png}
%   \endminipage\hfill
%   \minipage{0.32\textwidth}%
%     \includegraphics[width=\linewidth]{images/wave3.png}
%   \endminipage
%   \caption{Solutions to the wave equation for a given initial displacement.}
% \end{figure}

\begin{equation}
    u(r,\phi,t)=\frac{9}{4 \alpha_{03}}J_0(\alpha_{03}r)\sin(\alpha_{03} t)-\frac{1}{\alpha_{12}} J_1(\alpha_{12}r)\cos(\phi)\sin(\alpha_{12} t)+\frac{1}{5\alpha_{24}} J_2(\alpha_{24}r)\sin(2\phi)\sin(\alpha_{12} t)
\end{equation}

% \begin{figure}[!htb]
%     \minipage{0.32\textwidth}
%       \includegraphics[width=\linewidth]{images/waves1.png}
%     \endminipage\hfill
%     \minipage{0.32\textwidth}
%       \includegraphics[width=\linewidth]{images/waves2.png}
%     \endminipage\hfill
%     \minipage{0.32\textwidth}%
%       \includegraphics[width=\linewidth]{images/waves3.png}
%     \endminipage
%     \caption{Solutions to the wave equation for a given initial velocity.}
%     \end{figure}
\newpage
\section{References}
Asmar, N. H. (2002). \textit{Applied complex analysis with partial differential equations}. New Jersey: Prentice-Hall Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand differential equations in polar and spherical coordinate systems, as well as certain properties of Bessel functions.\\
\end{adjustwidth}
Farlow, S. J. (2016). \textit{Partial differential equations for scientists and engineers}. New York: Dover Publications, Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand the basic types of partial differential equations, as well as some methods for solving them in one dimension. It also introduced the physical interpretation of these equations.\\
\end{adjustwidth}
Fetter, A. L., \& Walecka, J. D. (2003). \textit{Theoretical mechanics of particles and continua}. Mineola, NY: Dover Publications. \\
\begin{adjustwidth}{2 cm}{}
This book was used to understand the physical meaning behind the wave and heat equation, as well as some theorems from Sturm-Liouville theory.\\
\end{adjustwidth}
Tolstov, G. P. (2009). \textit{Fourier series}. New York: Dover Publications, Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand basic properties of Fourier Series and Fourier-Bessel Series, as well as general methods of eigenfunction expansions.
\end{adjustwidth}
\end{document}

>>>>

\documentclass{article}
\usepackage{changepage}
\usepackage{float}
\usepackage{noah}

@@

\begin{equation}\label{selfadjoint}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width = 4.5 cm]{images/initialvalue.png}
  \caption{Initial distribution to be analyzed.}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width = 1.0\textwidth]{images/heat.png}
  \caption{Evolution of heat equation for initial distribution given.}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width = 1.0\textwidth]{images/wave.png}
  \caption{Evolution of wave equation for initial displacement given.}
\end{figure}

@@

    u(r,\phi,t)=\frac{9}{4 \alpha_{03}}J_0(\alpha_{03}r)\sin(\alpha_{03} t)-\frac{1}{\alpha_{12}} J_1(\alpha_{12}r)\cos(\phi)\sin(\alpha_{12} t)+\frac{1}{5\alpha_{24}} J_2(\alpha_{24}r)\sin(2\phi)\sin(\alpha_{12} t)

\begin{figure}[H]
  \centering
  \includegraphics[width = 1.0\textwidth]{images/wavevel.png}
  \caption{Evolution of wave equation for initial velocity given.}
\end{figure}
We can see that these solutions to the heat equation are oscillating about the steady state solution $u = 0$. We also see that for the solution where only the initial velocity is given, the amplitude is much lower than it is when the initial displacement is given. This is because we are dividing each term of the solution by its corresponding eigenvalue. This in turn means that the initial velocity does not impact the dynamics as much as the initial position does.
\section{Conclusion}
Through the method of eigenfunction expansions, solutions to the heat and wave equation are generated, which in turn allow us to analyze the dynamics of solutions. We see that the wave equation is the higher dimensional analogue to the simple harmonic oscillator, and the heat equation is the higher dimensional analogue to exponential decay. Instead of the solutions oscillating or decay towards and equilibrium point, they end up decaying to an equilibrium function, one that is harmonic.\\
It is important to reiterate that this method of solving functions by eigenfunction expansion is not unique to Laplace operator, and can be used for any operator that satisfies equation (\ref{selfadjoint}), known as a self-adjoint operator. This allows us to in principal solve nearly any linear-homogenous PDE. In fact this method has a direct analogy to methods for solving a linear system of ODEs. Traditionally, we attempt to find the eigenvalues and straight-line solutions to the equation, which corresponds to finding the normal modes of a differential operator.
\newpage
\section{References}
Asmar, N. H. (2002). \textit{Applied complex analysis with partial differential equations}. New Jersey: Prentice-Hall Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand differential equations in polar and spherical coordinate systems, as well as certain properties of Bessel functions.\\
\end{adjustwidth}
Farlow, S. J. (2016). \textit{Partial differential equations for scientists and engineers}. New York: Dover Publications, Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand the basic types of partial differential equations, as well as some methods for solving them in one dimension. It also introduced the physical interpretation of these equations.\\
\end{adjustwidth}
Fetter, A. L., \& Walecka, J. D. (2003). \textit{Theoretical mechanics of particles and continua}. Mineola, NY: Dover Publications. \\
\begin{adjustwidth}{2 cm}{}
This book was used to understand the physical meaning behind the wave and heat equation, as well as some theorems from Sturm-Liouville theory.\\
\end{adjustwidth}
Jackson, J. D. (1972). Classical electrodynamics (1st ed.). New York: Wiley.\\
\begin{adjustwidth}{2 cm}{}
This book was useful in introducing the method of eigenfunction expansions, Green's functions, and different types of boundary value problems.\\
\end{adjustwidth}
Tolstov, G. P. (2009). \textit{Fourier series}. New York: Dover Publications, Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand basic properties of Fourier Series and Fourier-Bessel Series, as well as general methods of eigenfunction expansions.
\end{adjustwidth}
\end{document}
